A Horrocks-type theorem for even orthogonal $K_2$

### 2020-03-31 dl-paper.tex
Fix 12

\documentclass[oneside, 8pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, mathtools, xfrac, microtype, nameref, thmtools}
\usepackage[breaklinks=false,unicode]{hyperref}
\usepackage[backend=biber, bibencoding=utf8, giveninits=true, citestyle=numeric-comp, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}
\usepackage[capitalize]{cleveref}
\usepackage[matrix,arrow,curve]{xy}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage[notref,notcite]{showkeys}

@@

\begin{df} \label{defP0}
Let $M$ be an ideal of a local ring $A$. Denote by $\widetilde{P}_{\alpha, M}(0)$ the subgroup of $\overline{\St}(\Phi, A[X], M[X])$ generated by the following five families of elements:
\begin{enumerate} 
 \item $z_{\beta}(Xf, \xi),\ \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi,\ f \in M[X],\ \xi \in A[X];$
 \item $z_{\beta}(f, X\xi),\ \alpha - \beta \in \Phi,\ f\in M[X],\ \xi\in A[X];$
 \item $z_{\beta}(f, \xi),\ \beta \perp \alpha,\ f\in M[X],\ \xi\in A[X];$
 \item $x_{-\alpha}(X^2f),\ f \in M[X];$
 \item $x_{\alpha}(f),\ f \in M[X]$. \end{enumerate}
In the sequel we continue to denote the elements of $M[X]$ (resp. $A[X]$) by $f$ (resp. $\xi$).
Notice also that by~\eqref{Zdef} one has $z_\beta(f, 0) = x_\beta(f)$.

We denote by $\widetilde{P}_{\alpha, M}(*)$ the subgroup of $\overline{\St}(\Phi, A[X], M[X])$ generated by $\widetilde{P}_{\alpha, M}(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$.
\end{df}
%We denote by $P_{\alpha, M}(0)$ the image of $\widetilde{P}_{\alpha, M}(0)$ under the map $j_+ \colon \St(\Phi, A[X]) \to \St(\Phi, A[X, X^{-1}])$. 
Almost always we will be using the above definition in the situation when $M$ is precisely the maximal ideal of $A$.
The only exception to this is~\cref{P0-conj-invariant} where the above subgroups are also used for $M=A$.

@@

Let $g = z_\beta(Xf, \xi)$ be a generator of type 1 such that $\alpha - \beta \in \Phi$. By~\cref{Zrels}\eqref{Z2} we get that \begin{equation} z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} =  x_{-\alpha} (- mX^2f\xi) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mX^2f) \cdot z_{\beta}(Xf, \xi). \label{eq3-1} \end{equation}
The first and the last factors in the right-hand side of the above equation are generators of type $4$ and $1$, respectively. It is easy to see that the second factor is also a generator of type 1, indeed:
\[ x_{\beta - \alpha} (N_{\beta,-\alpha} \cdot mX^2 f) = z_{\beta'}(Xf', 0)\text{ for } f' = N_{\beta,-\alpha} \cdot mX f\text{ and }\beta' = \beta - \alpha \text{ (clearly, $\alpha + \beta' \in \Phi$).}\]
Now consider the case when $g = z_\beta(f, X\xi)$ is a generator of type $2$. Again by~\cref{Zrels}\eqref{Z2}
\begin{equation}
z_{\beta}(f, X\xi) ^ {x_{-\alpha}(mX)} =  x_{-\alpha} (- mX^2f\xi ) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mXf) \cdot z_{\beta}(f, X\xi). \label{eq3-2} \end{equation}
It is easy to see that the expression in the right-hand side of~\eqref{eq3-2} is a product of generators of type 4, 1, 2.

@@

In the sequel we often use the following property of $S_\alpha(a, -)$.
\begin{lemma} \label{lem:Smult}
For $g_1 \in P_\alpha(m)$, $g_2 \in P_\alpha(*)$ one has
$S_\alpha(a, g_1\cdot g_2) = S_\alpha(a, g_1) \cdot S_\alpha\left(\tfrac{a}{1+am}, g_2\right).$
\end{lemma} \begin{proof} Since $g_2\in P_\alpha(m')$ for some $m' \in M$, we have that $g_1 \cdot g_2 \in P_\alpha(m + m')$. The assertion now follows from the definition of $S_\alpha(a, -)$:
\begin{multline*} S_\alpha(a, g_1 \cdot g_2) = x_\alpha(aX^{-1}) \cdot g_1 \cdot g_2 \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1+am+am'}\right) \cdot \{X, 1 + am + am'\} = \\ = x_\alpha(aX^{-1}) \cdot g_1 \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1+am}\right) \cdot x_\alpha\left(\tfrac{aX^{-1}}{1+am}\right) \cdot g_2 \cdot x_\alpha\left(-\frac{\tfrac{a}{1+am}\cdot X^{-1}}{1 + \tfrac{a}{1+am}\cdot m'}\right) \{X, (1+am)(1 + \tfrac{am'}{1+am})\} = \\ = S_\alpha(a, g_1) \cdot S_\alpha\left(\tfrac{a}{1+am}, g_2\right). \qedhere \end{multline*}
\end{proof}
It is clear that the restriction of the map $S_\alpha(a, g)$ to the subgroup $P_\alpha(0)$ coincides with the homomorphism of left conjugation by $x_\alpha(aX^{-1})$.
\begin{lemma}\label{P0_conj} The subgroup $P_\alpha(0)$ is stable under conjugation by $x_\alpha(aX^{-1})$ for arbitrary $a\in A$. \end{lemma}
\begin{proof} 
We need to verify that the conjugate by $x_\alpha(aX^{-1})$ to every generator $g$ from~\cref{defP0} lies in $P_\alpha(0)$. 
The assertion is clear for generators of type 3 and 5.
First of all, consider the case $\alpha + \beta \in \Phi$, which is only possible in the case of a generator of type 1:
\begin{equation}\label{eq:Zconj1} z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (- af\xi) \cdot x_{\alpha+\beta} (N_{\beta, \alpha}\cdot af) \cdot z_{\beta}(Xf, \xi). \end{equation}
Clearly, the expression in the right-hand side is a product of generators of type 5, 2, 1.

Now consider the case $\alpha - \beta \in \Phi$, which is possible for generators of type 1 and 2. By~\cref{Zrels} we obtain that
\begin{align} z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = &  x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot af\xi^2) \cdot z_{\beta}(Xf, \xi), \label{eq:Zconj2} \\
z_{\beta}(f, X\xi) ^ {x_{\alpha}(aX^{-1})} = & x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot aXf\xi^2) \cdot z_{\beta}(f, X\xi). \nonumber \end{align}
The expressions in the right-hand side are products of generators of type 5, 2, 1 and 5, 1, 2.

>>>>

\documentclass[oneside, 8pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, mathtools, xfrac, microtype, nameref, thmtools}
\usepackage[breaklinks=false,unicode]{hyperref}
\usepackage[backend=biber, bibencoding=utf8, giveninits=true, citestyle=numeric-comp, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}
\usepackage[capitalize]{cleveref}
\usepackage[matrix,arrow,curve]{xy}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage[notref,notcite]{showkeys}
\usepackage{MnSymbol}

@@

\begin{df} \label{defP0}
Let $M$ be an ideal of a local ring $A$. Denote by $\widetilde{P}_{\alpha, M}(0)$ the subgroup of $\overline{\St}(\Phi, A[X], M[X])$ generated by the following five families of elements:
\begin{enumerate} 
 \item $z_{\beta}(Xf, \xi),\ \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi,\ f \in M[X],\ \xi \in A[X];$
 \item $z_{\beta}(f, X\xi),\ \alpha - \beta \in \Phi,\ f\in M[X],\ \xi\in A[X];$
 \item $z_{\beta}(f, \xi),\ \beta \perp \alpha,\ f\in M[X],\ \xi\in A[X];$
 \item $x_{-\alpha}(X^2f),\ f \in M[X];$
 \item $x_{\alpha}(f),\ f \in M[X]$. \end{enumerate}

 We denote by $\widetilde{P}_{\alpha, M}(*)$ the subgroup of $\overline{\St}(\Phi, A[X], M[X])$ generated by $\widetilde{P}_{\alpha, M}(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$.
\end{df}

Almost always we will be using the above definition in the situation when $M$ is precisely the maximal ideal of $A$.
The only exception to this is~\cref{P0-conj-invariant} where the above subgroups are also used for $M=A$.
\begin{rem} \label{rem:recognition}
  From the above definition it follows that an element $z_\gamma(f, \xi)$ is a generator of type 1 provided $X \divides f$, $\gamma\neq\pm\alpha$ and $\gamma$ is not orthogonal to $\alpha$.
  Since $x_\gamma(f) = z_\gamma(f, 0)$, it follows that under the same assumptions $x_{\gamma}(f)$ is also a generator of type 1.
  In particular, if $\beta, \alpha - \beta \in \Phi$ then $x_{\pm(\alpha-\beta)}(Xf)$ is a generator of type 1.

  Similarly, if $\gamma - \alpha \in \Phi$ and $X \divides \xi$ then $z_\gamma(f, \xi)$ is a generator of type 2.
  In particular, if $\beta, \alpha + \beta \in \Phi$ then $x_{\alpha+\beta}(f)$ is a generator of type 2.
\end{rem}

@@

Let $\beta\in\Phi$ be such that $\alpha - \beta \in \Phi$ and $g = z_\beta(Xf, \xi)$ be a generator of type 1. By~\cref{Zrels}\eqref{Z2} we get that \begin{equation} z_{\beta}(Xf, \xi) ^ {x_{-\alpha}(mX)} =  x_{-\alpha} (- mX^2f\xi) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mX^2f) \cdot z_{\beta}(Xf, \xi). \label{eq3-1} \end{equation}
The expression in the right-hand side is a product of generators of type 4, 1, 1 (see~\cref{rem:recognition}).
Now consider the case when $g = z_\beta(f, X\xi)$ is a generator of type $2$ (so that $\alpha-\beta\in\Phi$). Again by~\cref{Zrels}\eqref{Z2}
\begin{equation}
z_{\beta}(f, X\xi) ^ {x_{-\alpha}(mX)} =  x_{-\alpha} (- mX^2f\xi ) \cdot x_{\beta-\alpha} (N_{\beta, -\alpha}\cdot mXf) \cdot z_{\beta}(f, X\xi). \label{eq3-2} \end{equation}
It is easy to see that the expression in the right-hand side of~\eqref{eq3-2} is a product of generators of type 4, 1, 2.

@@

In the sequel we often use the following property of $S_\alpha(a, -)$.
\begin{lemma} \label{lem:Smult}
For $g_1 \in P_\alpha(m)$, $g_2 \in P_\alpha(*)$ one has
$S_\alpha(a, g_1\cdot g_2) = S_\alpha(a, g_1) \cdot S_\alpha\left(\tfrac{a}{1+am}, g_2\right).$
\end{lemma} \begin{proof} Assume $g_2\in P_\alpha(m')$ for some $m' \in M$, so that $g_1 \cdot g_2 \in P_\alpha(m + m')$. The assertion now follows from the definition of $S_\alpha(a, -)$:
\begin{multline*} S_\alpha(a, g_1 \cdot g_2) = x_\alpha(aX^{-1}) \cdot g_1 \cdot g_2 \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1+am+am'}\right) \cdot \{X, 1 + am + am'\} = \\ = x_\alpha(aX^{-1}) \cdot g_1 \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1+am}\right) \cdot x_\alpha\left(\tfrac{aX^{-1}}{1+am}\right) \cdot g_2 \cdot x_\alpha\left(-\frac{\tfrac{a}{1+am}\cdot X^{-1}}{1 + \tfrac{a}{1+am}\cdot m'}\right) \{X, (1+am)(1 + \tfrac{am'}{1+am})\} = \\ = S_\alpha(a, g_1) \cdot S_\alpha\left(\tfrac{a}{1+am}, g_2\right). \qedhere \end{multline*}
\end{proof}
It is clear that the restriction of the map $S_\alpha(a, g)$ to the subgroup $P_\alpha(0)$ coincides with the homomorphism of left conjugation by $x_\alpha(aX^{-1})$.
\begin{lemma}\label{P0_conj} The subgroup $P_\alpha(0)$ is stable under conjugation by $x_\alpha(aX^{-1})$ for arbitrary $a\in A$. \end{lemma}
\begin{proof} 
Since $P_\alpha(0)$ has the same generators as $\widetilde{P}_{\alpha, M}(0)$ we need to verify that the conjugate by $x_\alpha(aX^{-1})$ to every generator from~\cref{defP0} lies in $P_\alpha(0)$. 
The assertion is clear for the generators of type 3 and 5.

Let $\beta$ be such that $\alpha + \beta \in \Phi$ and $g = z_\beta(Xf, \xi)$ be a generator of type $1$.
By~\cref{Zrels}\eqref{Z2}
\begin{equation}\label{eq:Zconj1} z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (- af\xi) \cdot x_{\alpha+\beta} (N_{\beta, \alpha}\cdot af) \cdot z_{\beta}(Xf, \xi), \end{equation}
and the latter expression a product of generators of type 5, 2, 1 (see~\cref{rem:recognition}).

Let $\beta$ be such that $\alpha - \beta \in \Phi$ and let $g = z_{\beta}(Xf, \xi)$ be a generator of type $1$. From~\cref{Zrels}~\eqref{Z3} we obtain that
\begin{equation} z_{\beta}(Xf, \xi) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot af\xi^2) \cdot z_{\beta}(Xf, \xi), \label{eq:Zconj2} \end{equation}
and the expression in the right-hand side is a product of generators of type 5, 2, 1.
Now let $g = z_{\beta}(f, X\xi)$ be a generator of type 2 (for $\beta$ as above). Again by~\cref{Zrels}~\ref{Z3}
\begin{equation} z_{\beta}(f, X\xi) ^ {x_{\alpha}(aX^{-1})} = x_{\alpha} (af\xi) \cdot x_{\alpha-\beta} (N_{\beta,-\alpha}\cdot aXf\xi^2) \cdot z_{\beta}(f, X\xi).
\end{equation}
and the expression in the right-hand side are products of generators of type 5, 1, 2.
